Differential expression analysis of reprocessed AMPAD RNASeq data (All 7 brain regions with CQN normlaisation and BIC based model selection with study regressed apriori)

### 2017-08-09 code/Rmd/ALL_REPROCESSED.Rmd
Minor changes

  x[x < (mn-3*std.dev)] = x[ind.low]
  x[x > (mn+3*std.dev)] = x[ind.high]
  return(x)
}) %>% t
CQN.GENE_EXPRESSION$E = LOG.CPM
```

@@

>>>>

  x[x < (mn-3*std.dev)] = x[ind.low]
  x[x > (mn+3*std.dev)] = x[ind.high]
  return(x)
}) %>% t
CQN.GENE_EXPRESSION$E = LOG.CPM
NEW.COUNTS = (2^LOG.CPM) * t(replicate(dim(LOG.CPM)[1], LIB.SIZE))/1e6
```

@@

# Store filtered counts
NEW.COUNTS %>%
  rownameToFirstColumn('Gene.ID') %>%
  write.table(file = 'AMP_nCounts.tsv', sep = '\t', row.names=F, quote=F)
NCOUNT_OBJ = File('AMP_nCounts.tsv', name = 'Counts (estimated)', parentId = CODE$properties$id)
annotations(NCOUNT_OBJ) = annotations(CODE)
annotations(NCOUNT_OBJ)$dataSubType = 'estimatedCounts'
NCOUNT_OBJ = synStore(NCOUNT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

### 2017-08-09 code/Rmd/ALL_REPROCESSED_v2.Rmd
Minor changes

### Model Identification
```{r fxn}
# Function to fit a mixed effect model and return BIC
fitMixedEffectModel <- function(varsToFit, groupingVar, lcpm, voomWeights, covars){

@@

  return(varsSelected)
}
```
#### Regress out Study 
```{r regress.study}
# Get design matrix
DESIGN.MAT = getDesignMatrix(COVARIATES[colnames(NEW.COUNTS), 'Study', drop=F], Intercept = F)
DESIGN.MAT = DESIGN.MAT$design[,linColumnFinder(DESIGN.MAT$design)$indepCols]

@@

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=BrainRegion, shape=Study, size=RIN))
p <- p + theme_bw() + theme(legend.position="right")
p
```
#### Phase I (clinical, ancestry and sample specific technical variables)
```{r phase1}
# Fit a base model with the following covariates
adjust.covars = c('BrainRegion.Diagnosis')
baseModelBIC = fitMixedEffectModel(adjust.covars, 
                                   c('individualIdentifier'), 
                                   RESIDUAL.EXPR, 
                                   VOOM.WEIGHTS$weights, 
                                   COVARIATES) %>%
  dplyr::rename(baseModel = BIC)

@@

#### Phase II (batch effects)
```{r phase2}
# Fit a base model with the following covariates
baseModelBIC = fitMixedEffectModel(adjust.covars, 
                                   c('individualIdentifier'), 
                                   RESIDUAL.EXPR, 
                                   VOOM.WEIGHTS$weights, 
                                   COVARIATES) %>%
  dplyr::rename(baseModel = BIC)

@@

#### Phase III (RNASeq alignment specific covariates)
```{r phase3}
# Fit a base model with the following covariates
baseModelBIC = fitMixedEffectModel(adjust.covars, 
                                   c('individualIdentifier'), 
                                   RESIDUAL.EXPR, 
                                   VOOM.WEIGHTS$weights, 
                                   COVARIATES) %>%
  dplyr::rename(baseModel = BIC)

@@

# Estimate consensus correlation coefficient
correlation <- parallelDuplicateCorrelation(VOOM.GENE_EXPRESSION, block = COVARIATES$Individual_ID)

@@

# Fit linear model using new weights and new design
VOOM.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION$E
ADJUSTED.FIT = lmFit(VOOM.GENE_EXPRESSION)

# Residuals after normalisation
RESIDUAL.GENE_EXPRESSION = residuals.MArrayLM(ADJUSTED.FIT, CQN.GENE_EXPRESSION$E)

@@

# Calculate correlation values of random effects
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$E
correlation = parallelDuplicateCorrelation(VOOM.WEIGHTS, block = COVARIATES$individualIdentifier)

@@

# Code
CODE <- Folder(name = "ALL Reprocessed RNASeq (Normalised)", parentId = parentId)
annotations(CODE) = all.annotations
CODE <- synStore(CODE)
synChangeEntityACL(CODE@properties$id, new.racl)

>>>>

### Model Identification
```{r fxn, eval = F}
# Function to fit a mixed effect model and return BIC
fitMixedEffectModel <- function(varsToFit, groupingVar, lcpm, voomWeights, covars){

@@

  return(varsSelected)
}
```
#### Regress out Study 
```{r regress.study}
# Get design matrix
DESIGN.MAT = getDesignMatrix(COVARIATES[colnames(NEW.COUNTS), 'Study', drop=F], Intercept = T)
DESIGN.MAT = DESIGN.MAT$design[,linColumnFinder(DESIGN.MAT$design)$indepCols]

@@

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=BrainRegion, shape=Study, size=RIN))
p <- p + theme_bw() + theme(legend.position="right")
p

adjust.covars = c('BrainRegion.Diagnosis', 'RIN', 'RIN2', 'Gender', 'PMI', 'AOD', 'Batch', 'PCT_RIBOSOMAL_BASES', 
                  'PCT_INTRONIC_BASES', 'PCT_CODING_BASES', 'PCT_PF_READS_ALIGNED', 'PCT_INTERGENIC_BASES')
```
#### Phase I (clinical, ancestry and sample specific technical variables)
```{r phase1, eval = F}
# Fit a base model with the following covariates
adjust.covars = c('BrainRegion.Diagnosis')
baseModelBIC = fitMixedEffectModel(adjust.covars, 
                                   c('individualIdentifier'), 
                                   RESIDUAL.EXPR, 
                                   VOOM.WEIGHTS$weights, 
                                   COVARIATES) %>%
  dplyr::rename(baseModel = BIC)

@@

#### Phase II (batch effects)
```{r phase2, eval = F}
# Fit a base model with the following covariates
baseModelBIC = fitMixedEffectModel(adjust.covars, 
                                   c('individualIdentifier'), 
                                   RESIDUAL.EXPR, 
                                   VOOM.WEIGHTS$weights, 
                                   COVARIATES) %>%
  dplyr::rename(baseModel = BIC)

@@

#### Phase III (RNASeq alignment specific covariates)
```{r phase3, eval = F}
# Fit a base model with the following covariates
baseModelBIC = fitMixedEffectModel(adjust.covars, 
                                   c('individualIdentifier'), 
                                   RESIDUAL.EXPR, 
                                   VOOM.WEIGHTS$weights, 
                                   COVARIATES) %>%
  dplyr::rename(baseModel = BIC)

@@

# Estimate consensus correlation coefficient
# correlation <- parallelDuplicateCorrelation(VOOM.GENE_EXPRESSION, block = COVARIATES$Individual_ID)
correlation = list()
correlation$cor = 0.45

@@

# Fit linear model using new weights and new design
VOOM.GENE_EXPRESSION$E = RESIDUAL.EXPR
ADJUSTED.FIT = lmFit(VOOM.GENE_EXPRESSION)

# Residuals after normalisation
RESIDUAL.GENE_EXPRESSION = residuals.MArrayLM(ADJUSTED.FIT, VOOM.GENE_EXPRESSION$E)

@@

# Calculate correlation values of random effects
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$E
# correlation = parallelDuplicateCorrelation(VOOM.WEIGHTS, block = COVARIATES$individualIdentifier)

@@

# Code
CODE <- Folder(name = "ALL Reprocessed RNASeq (Normalised) - study effects removed", parentId = parentId)
annotations(CODE) = all.annotations
CODE <- synStore(CODE)
synChangeEntityACL(CODE@properties$id, new.racl)
